\documentclass[12pt]{article}
\usepackage[letterpaper, margin=2cm]{geometry}

\begin{document}
  I am writing this statement to request of waiver of the registration fees for
  the 2021 SIAM Computation Sciences and Engineering conference.
  I am requesting this waiver as I am a Ph.D. student in my final year of study,
  and any assistance in attending this conference is very helpful and 
  appreciated.

  I am also very interested in attending this conference.
  I am interested in attending this conference as it will be an excellent source
  of information related to my field of study, computational PDEs.
  Specifically I am interested in attending MS152, Advanced Developement on 
  Computational Methods for Kinetic Theory, MS283, High-Performance PDE Solvers,
  and MS299, A Deep Dive into Shallow Water.
  It will also be an excellent opportunity to network as I am in the middle of
  searching for jobs.

  I am also scheduled to give a talk at this conference.
  My presentation is titled \emph{Nonconservative Discontinuous Galerkin Method 
  for Generalized Shallow Water Equations on a Sphere}.
  It is coauthored by my advisor, James Rossmanith, of Iowa State University.
  Originally this presentation was submitted as a contributed talk, but has now
  been scheduled as part of minisymposium 311, Exascale Applications for 
  High-Order PDE Solvers.
  The organizers of this minisymposium are MiSun Min and Paul Fischer.

  \vspace{0.5cm}

  \noindent Best Regards,

  \noindent Caleb Logemann

\end{document}
