%\documentclass[draft]{elsarticle}
\documentclass{elsarticle}
\usepackage[margin=2.5cm]{geometry}

\input{styles.tex}
\def\B{{\bf B}}
\def\half{\frac{1}{2}}
\def\En{{\mathcal{E}}}
\def\E{{\bf E}}
\def\J{{\bf J}}
\def\ft{\tilde{f}}
\def\tti{\tilde{t}}
\def\xt{\tilde{x}}
\def\vt{\tilde{v}}
\def\Et{\tilde{E}}
\def\rhot{\tilde{\rho}}
\def\Flux{{\mathcal{F}}}

\newcommand{\M}[1]{M_{#1}}

\usepackage{algorithm,algorithmic}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
%\theoremstyle{remark}
%\newtheorem*{rmk}{Remark}
\newtheorem{remark}{Remark}

\journal{J. Comput. Appl. Math.}

\begin{document}

\begin{frontmatter}

%\title{Asymptotic-Preserving Quadrature-Based Moment-Closure Methods
%for the Vlasov-Poisson-\\ Fokker-Planck System in the High-Field Limit}

\title{An Implicit-Explicit Discontinuous Galerkin Scheme using a Newton-Free Picard Iteration for a Thin-Film Model}

\author[author1]{Caleb Logemann}
\ead{logemann@iastate.edu}

\author[author1]{James A. Rossmanith\fnref{labc}}
\ead{rossmani@iastate.edu}

\address[author1]{Iowa State University, Department of Mathematics,
396 Carver Hall, Ames, IA 50011, USA}

%\address[author2]{Iowa State University, Department of Mathematics,
%396 Carver Hall, Ames, IA 50011, USA}



\fntext[labc]{Corresponding author}

\begin{abstract}
  This paper provides a high-order numerical scheme for solving thin-film models of the
  form \(q_t + \p{q^2 - q^3}_x = -\p{q^3 q_{xxx}}_x\).
  The second term in this equation is of nonlinear hyperbolic type, while the right-hand
  side is of nonlinear parabolic type.
  The nonlinear hyperbolic term is discretized with the standard modal discontinuous
  Galerkin method, and the nonlinear parabolic term is discretized with the local
  discontinuous Galerkin method.
  Propagation in time is done with an implicit-explicit Runge-Kutta scheme so as to
  allow for larger time steps.
  The timestep restriction for these methods is determined by the hyperbolic wavespeed
  restriction, and is not limited by the nonlinear parabolic term.
  % TODO: Describe as a linearization or state that only 1 Picard iteration is needed to converge
  A novel aspect of this method is that the resulting nonlinear algebraic equations
  are solved with a Newton-free iteration with a Picard iteration.
  The number of iterations required to converge is less than or equal to the estimated
  order of the method.
  We have demonstrated with this method up to third order convergence.
\end{abstract}

\begin{keyword}
Discontinuous Galerkin; High-Order Schemes; Implicit-Explicit; Newtwon-Free; Picard Iteration; Thin-Film Models
\end{keyword}

\end{frontmatter}

\input{introduction.tex}
\input{equations.tex}
\input{methods.tex}
\input{results.tex}
\input{conclusions.tex}

\bigskip

\noindent
{\bf Acknowledgements.}
This work was supported in part by NSF grant DMS--1620128.

\bibliographystyle{abbrv}
\bibliography{bibliography/refs.bib}

\end{document}
